### 2022-05-31 paper.Rmd
shortened abstract

    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity. Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the k-means criterion. Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity. While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments, and covariance structure. To minimize differences in variance, each variable is duplicated by computing the squared deviation between the original values and the variable's mean. Because the variance is defined as the expected value of the squared deviations from the mean, appending these duplicated variables to the data set---and then applying standard k-means anticlustering---suffices to equalize the variance between groups. Following a similar logic, differences in skewness, kurtosis and covariances can be minimized. Thus, while k-plus is a separate criterion for anticlustering, it can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mean attribute values as well as the variance and higher order moments, usually with no noticable sacrifice to either criterion. The k-plus extension is therefore preferred over classical k-means anticlustering for optimizing between-group similarity. Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).

>>>>

    Anticlustering refers to a collection of methods that partition data sets into disjoint groups while aiming for high between-group similarity and high within-group heterogeneity. Anticlustering is usually approached by maximzing instead of minimizing a clustering objective such as the k-means criterion. Introducing k-plus, this paper presents an extension of k-means to maximize between-group similarity. While k-means anticlustering minimizes between-group-differences only with regard to the mean of attribute values, k-plus anticlustering can be used to minimize between-group-differences with regard to the variance, higher order moments like skewness and kurtosis, and even covariance structure. It is shown that all k-plus variations can be reduced to an augmentation of the input data while the original k-means objective remains unchanged. A computer simulation and practical examples on real data show that k-plus anticlustering achieves high similarity with regard to mutliple distribution characteristics with little sacrifice differences in means. The k-plus extension is therefore preferred over classical k-means anticlustering when optimizing between-group similarity. Example are given on how k-plus anticlustering can be applied using the open source R package `anticlust`, which is freely available via CRAN (https://cran.r-project.org/package=anticlust) and Github (https://github.com/m-Py/anticlust).
